Cox Proportional Hazards Model

### 2025-02-10 index.qmd
fixed article summaries (forget self-contained)
Add intro draft to index.qmd

---
title: "Cox Proportional Hazards Model"
subtitle: "Temp"
author: "Jayme Reed (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

::: callout-important
**Remember:** Your goal is to make your audience understand and care
about your findings. By crafting a compelling story, you can effectively
communicate the value of your data science project.

Carefully read this template since it has instructions and tips to
writing!
:::

@@

*This is an introduction to ..... regression, which is a non-parametric
estimator that estimates the conditional expectation of two variables
which is random. The goal of a kernel regression is to discover the
non-linear relationship between two random variables. To discover the
non-linear relationship, kernel estimator or kernel smoothing is the
main method to estimate the curve for non-parametric statistics. In
kernel estimator, weight function is known as kernel function
[@efr2008]. Cite this paper [@bro2014principal]. The GEE [@wang2014].
The PCA [@daffertshofer2004pca]*

@@

>>>>

---
title: "Cox Proportional Hazards Model"
subtitle: "Temp"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

@@

*This is an introduction to ..... regression, which is a non-parametric
estimator that estimates the conditional expectation of two variables
which is random. The goal of a kernel regression is to discover the
non-linear relationship between two random variables. To discover the
non-linear relationship, kernel estimator or kernel smoothing is the
main method to estimate the curve for non-parametric statistics. In
kernel estimator, weight function is known as kernel function
[@efr2008]. Cite this paper [@bro2014principal]. The goat study*
[@ziadi2023]*. The insurance[@zapletal2021]*

@@

## Introduction (draft)

Analyzing survival data is challenging for traditional machine learning models because they struggle to capture the time-to-event nature of the data.Survival data typically contains challenges difficult to account for because the datasets may be “censored” meaning that the event hasn’t had the opportunity to occur yet by the end of the study period. Also, traditional methods aren’t able to effectively model time-to-event data because they treat time as fixed values opposed to treating the values as time leading to less accurate results.

Cox proportional hazards models are commonly used to interpret, analyze, and examine survival data. These models specialize in time-to-event predictions and can account for multiple variables and their effects on survival. Cox proportional hazards models are commonly used in the health field while predicting patient survivability based on symptoms and medical history, the financial field to predict bank failure, and mechanical studies to predict failure of machines. Cox proportional hazards models are popular due to their ability to incorporate multiple covariates and they have the advantage of being semi-parametric which allows more flexibility as compared to other models.

In this paper, we will be examining the prediction of remaining useful life (RUL)  for NASA turbofan jet engines. The lifespan of a jet engine is important as they cost thousands of dollars to maintain and produce. The ability to predict the remaining useful lifespan of these engines will allow for better judgments on the finances needed to maintain or repair them as well as alert teams that an engine may not have the RUL for a mission they would like to use it for.

We will first generate the summary statistics for this dataset as well as run linear and logistic regression tests. We will then test our assumptions which include the proportional hazards assumption, the linearity between covariates and the log hazard, that each data is independent of each other and there is no omitted data. A model will be generated using the survival package from R. We will then test the model and make adjustments as needed before generating our predictions.
